---
title: "CA_Drug_Study"
author: "Tyson Jones"
date: "2024-01-18"
output: html_document
---

```{r setup, include=FALSE}
library("tidyverse")
library("readxl")
library("moderndive")
library("corrplot")
id = "19Jo2mO7wHqX_g5dWdSSSV6hh1YN2TAAo"
CA_drug <- read.csv(sprintf("https://docs.google.com/uc?id=%s&export=download", id))
rm(id)

library(googledrive)



```
## First we will select users of drugs that affect NE levels.
### We are also going to determine chronic use by multiplying years * frequency.
### We will also separate users from nonusers for this set of analyses.
```{r}
#Splitting PRMQ by P and R
P_columns <- c("PRMQ_1","PRMQ_3","PRMQ_5","PRMQ_7","PRMQ_10","PRMQ_12","PRMQ_14","PRMQ_16")  

CA_drug <- CA_drug %>%
  mutate(Avg_PMQ = rowMeans(select(., all_of(P_columns)), na.rm = TRUE),.after = Avg_PRMQ)

R_columns <- c("PRMQ_2","PRMQ_4","PRMQ_6","PRMQ_8","PRMQ_9","PRMQ_11","PRMQ_13","PRMQ_15") 

CA_drug <- CA_drug %>%
  mutate(Avg_RMQ = rowMeans(select(., all_of(R_columns)), na.rm = TRUE),.after = Avg_PMQ)

rm(P_columns,R_columns)
```

```{r Selection, message=FALSE}
CA_nonusers <- CA_drug %>% 
  filter(user_nonuser == 1)

CA_users <- CA_drug %>%
  filter(user_nonuser == 2) %>% 
  mutate(NE_yn = if_else(MDMA_yn == 2 | coc_yn == 2 | meth_yn == 2,2,1,)) %>% 
  mutate(NE_yn = factor(NE_yn, levels = c(1, 2))) %>% 
  mutate(Avg_use = Avg_freq * Avg_totalyrs, .after = Avg_freq)
  
```

# Plot Time!

## As we can see, MDMA use in this group is moderately positively correlated with Cocaine and Meth use.

```{r Corr Plot, echo=FALSE}
CA_corr_yn <- CA_users %>%
  select(24:32) %>%
  cor()
  col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
  corrplot(CA_corr_yn, method = "shade", shade.col = NA, tl.col = "black", tl.srt = 45,col = col(200), addCoef.col = "black", cl.pos = "n", order = "AOE")

```

## This plot shows some funky-ness in PRMQ. Let's see if it's statistically significant

```{r , echo=FALSE}
CA_users %>%
  ggplot(aes(x= Avg_PRMQ, fill = NE_yn)) +
  geom_histogram(binwidth = .5) +
  facet_wrap(~ NE_yn, scales = "free_y")

```

```{r}
# Check the variances
variances <- tapply(CA_users$Avg_PRMQ, CA_users$NE_yn, var, na.rm = TRUE)
print(variances)
rm(variances)
# Looks good now let's run a t test
NE_ttest <- t.test(CA_users$Avg_PRMQ ~ CA_users$NE_yn)
print(NE_ttest)
rm(NE_ttest)
```

### Seems significant, now let's ask Mr Cohen what he thinks.

```{r}
NE_user <- CA_users$Avg_PRMQ[CA_users$NE_yn == 2]
NE_nonuser <- CA_users$Avg_PRMQ[CA_users$NE_yn == 1]

cohens_d <- (mean(NE_user) - mean(NE_nonuser)) / sqrt((sd(NE_user)^2 + sd(NE_nonuser)^2) / 2)

print(cohens_d)
```

## Rule of thumb says this is a small-moderate effect, which is still exciting!


## Now to check to see if "Norepinephrine users" had any other major differences on their scores.
 
```{r echo=FALSE}
perform_analysis <- function(data, variable_name, group_variable) {
  # Check the variances
  variances <- tapply(data[[variable_name]], data[[group_variable]], var, na.rm = TRUE)
  cat("Variances for", variable_name, ":\n", variances, "\n\n")

  # Run a t-test
  t_test_result <- t.test(data[[variable_name]] ~ data[[group_variable]])
  cat("T-test result for", variable_name, ":\n")
  print(t_test_result)
  cat("\n")

  # Calculate Cohen's d
  group1_data <- data[[variable_name]][data[[group_variable]] == levels(data[[group_variable]])[1]]
  group2_data <- data[[variable_name]][data[[group_variable]] == levels(data[[group_variable]])[2]]
  cohens_d <- (mean(group1_data) - mean(group2_data)) / sqrt((sd(group1_data)^2 + sd(group2_data)^2) / 2)
  cat("Cohen's d for", variable_name, ":\n", cohens_d, "\n\n")
}

perform_analysis(CA_users, "Avg_PRMQ", "NE_yn")

perform_analysis(CA_users, "Avg_MI_EO", "NE_yn")

perform_analysis(CA_users, "Avg_MI_EC", "NE_yn")

perform_analysis(CA_users, "Avg_RTDT", "NE_yn")

perform_analysis(CA_users, "Avg_SC", "NE_yn")

perform_analysis(CA_users, "Avg_EQ", "NE_yn")

rm(cohens_d,NE_nonuser,NE_user)

```

## It looks like PRMQ and RTDT are the only two variables with statistically significant (p < .001 and p < .01 respectively) differences. Both have an effect size ~ .3 which is pretty solid!
## Let's see a graph for RTDT
```{r echo=FALSE}
CA_users %>%
  ggplot(aes(x= Avg_RTDT, fill = NE_yn)) +
  geom_histogram(binwidth = .5) +
  facet_wrap(~ NE_yn, scales = "free_y")

```

## Are there any differences between Norep users and other users?

```{r}
CA_users2 <- CA_users %>% 
  group_by(NE_yn)

CA_users2 %>% 
  ggplot(aes(x=Num_drugs, fill = NE_yn)) +
  geom_bar() + 
  facet_wrap(~ NE_yn, scales = "free")

```

Does safety of drug use affect the AC measures?

```{r Safety}

# File ID of the file you want to download
file_id <- "1_JoIgShPFobwutggd1DS78142vuDPpdD"
drive_download(as_id(file_id), overwrite = T)

CA_Safety <-

CA_Safety <- CA_Safety %>% 
  mutate_all(as.numeric) %>% 
  left_join(CA_users, CA_Safety, by = "IDNUM")

# Selecting all the drug risk questions
columns_to_sum <- c(2,4,6,8,10,12,14,16,18)  

CA_Safety <- CA_Safety %>%
  mutate(Abuse_Risk = rowSums(select(., columns_to_sum), na.rm = TRUE)) 

rm(columns_to_sum)

CA_Safety %>% 
  ggplot(aes(x = Abuse_Risk, y = Avg_PRMQ)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ NE_yn, scales = "free_x") 


Safety_model <- lm(CA_Safety$Avg_PRMQ ~ CA_Safety$Abuse_Risk)
get_regression_table(Safety_model)

correlation <- cor.test(CA_Safety$Avg_PRMQ, CA_Safety$Abuse_Risk)
correlation
```

```{r}
# Z test for users vs nonusers PRMQ
Z_StatUsers <- ((mean(CA_users$Avg_PRMQ) - mean(CA_nonusers$Avg_PRMQ)) /
             sd(CA_users$Avg_PRMQ / sqrt(nrow(CA_users))))
pt(Z_StatUsers, (nrow(CA_users)-1),lower.tail = F)

```

```{r Z Test}
# Z test for NE users vs NE nonusers PRMQ
CA_NE_use <- CA_users %>% 
  filter(NE_yn == 2)
CA_NE_non <- CA_users %>% 
  filter(NE_yn == 1)
#Not significant really

Z_StatNE <- ((mean(CA_NE_use$Avg_PRMQ) - mean(CA_NE_non$Avg_PRMQ)) /
             sd(CA_NE_use$Avg_PRMQ / sqrt(nrow(CA_NE_use))))
NE_p <- pt(Z_StatNE, (nrow(CA_NE_use)-1), lower.tail = F)
# significant!

# Z tests for PMQ and RMQ for NE users vs NE nonusers
Z_StatPMQ <- ((mean(CA_NE_use$Avg_PMQ) - mean(CA_NE_non$Avg_PMQ)) /
             sd(CA_NE_use$Avg_PMQ / sqrt(nrow(CA_NE_use))))
PMQ_p <- pt(Z_StatPMQ, (nrow(CA_NE_use)-1), lower.tail = F)
PMQ_p
# significant!

Z_StatRMQ <- ((mean(CA_NE_use$Avg_RMQ) - mean(CA_NE_non$Avg_RMQ)) /
             sd(CA_NE_use$Avg_RMQ / sqrt(nrow(CA_NE_use))))
RMQ_p <- pt(Z_StatRMQ, (nrow(CA_NE_use)-1), lower.tail = F)
RMQ_p
# significant!

# Z test for risk factors between NE users and NE nonusers
CA_Risk_NEuse <- CA_Safety %>% 
  filter(NE_yn == 2)
CA_Risk_NEnon <- CA_Safety %>% 
  filter(NE_yn == 1)

Z_StatRisk <- ((mean(CA_Risk_NEuse$Abuse_Risk) - mean(CA_Risk_NEnon$Abuse_Risk)) / sd(CA_Risk_NEuse$Abuse_Risk / sqrt(nrow(CA_Risk_NEuse))))

risk_p <- pt(Z_StatRisk, (nrow(CA_Risk_NEuse)-1), lower.tail = F)
```

```{r Hist PMQ + RMQ + Risk}
CA_users %>% 
  ggplot(aes(x = Avg_PMQ, fill = NE_yn)) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30, aes(y = ..density..)) +
  scale_fill_manual(values = c("blue", "red")) +
  theme_minimal()

CA_users %>% 
  ggplot(aes(x = Avg_RMQ, fill = NE_yn)) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 25, aes(y = ..density..)) +
  scale_fill_manual(values = c("blue", "red")) +
  theme_minimal()

CA_Safety_filtered <- CA_Safety[complete.cases(CA_Safety$NE_yn), ]

CA_Safety_filtered %>% 
  ggplot(aes(x = Abuse_Risk, fill = NE_yn)) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30, aes(y = ..density..)) +
  theme_minimal()
rm (CA_Safety_filtered)

```

```{r Abuse_Risk Regression}

abuse_model <- lm(Avg_PRMQ ~ Abuse_Risk + num_drugs, data = CA_users)

```

